---
title: "NU tet pos omicron analysis - hetero and haplo responses included - NU"
author: "Rebecca Payne"
date: "17/01/2023"
output: html_document
---
Analysis of the dominant antigen specific responses from three batches. Samples are either stained for EBV-APC, or Spike/non Spike- PE responses. 

Fcs pre-gated on live, CD3+CD8+Tetramer+ events

Premessa processing removed unwanted parameters including CD3, CD8, CD4 and Tetramers.

A cut-off of 0.01% or less than 10 events applied during this script and added to the metadata file.

repeat to removeAz sample 10065 - only Pfizer

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(readxl) 
library(CATALYST)
library(cowplot)
library(flowCore)
library(scater)
library(SingleCellExperiment)
library(openxlsx)
library(ggpubr)
```


### Create SCE and use FLOWSom to identify monocytes
Load  FCS files (processed), metadata, panel and create flowset

```{r}
md <- "metadata_aurora_test_omicron.xlsx" 
md <- read_excel(md)       
head(data.frame(md)) 
```


```{r}
fs <- read.flowSet(md$file_name, transformation = F, truncate_max_range = F)
```

```{r}
panel <- "Panel_aurora_test.xlsx" 
panel <- read_excel(panel)                                         
head(data.frame(panel),30) 
```


```{r}
all(panel$fcs_colname %in% colnames(fs))
```

```{r}
setdiff(fs@frames$SIREN10004_EOI_EBV_J.fcs@parameters@data$desc,panel$antigen)
```


```{r}
md$infection <- factor(md$infection, levels = c("pre_omicron", "post_omicron"))      

md$batch <- factor(md$batch, levels = c("batch1", "batch2","batch3" ))

md$antigen <- factor(md$antigen, levels = c("Spike", "Non_Spike", "EBV" ))

md$hla <- factor(md$hla, levels = c("A0201", "A0301", "B0702", "A0101"))

md$serostatus <- factor(md$serostatus, levels = c("previously_infected", "naive"))

md$cut_off_10 <- factor(md$cut_off_10, levels = c("yes", "no"))

md$cut_off_40 <- factor(md$cut_off_40, levels = c("yes", "no"))

md$condition <- factor(md$condition, levels = c("pre_omicron_spike_naive", "post_omicron_spike_naive", "pre_omicron_spike_previously_infected", "post_omicron_spike_previously_infected", "post_omicron_nonspike_naive", "pre_omicron_nonspike_previously_infected", "post_omicron_nonspike_previously_infected",  "pre_omicron_EBV", "post_omicron_EBV"))


md$sample_id <- factor(md$sample_id, levels = md$sample_id[order(md$condition)])  
```


```{asis, echo = TRUE}
# see ?prepdata for notes on the sce. Aurora flow analysis needs a different cofactor applied. We have chosen 150. Also need to specify FACE = TRUE for flow data.
```

```{r}
sce <- prepData(fs, panel, md, transform = TRUE, cofactor = 150, FACS = TRUE, features = panel$fcs_colname, md_cols = list(file = "file_name", id = "sample_id", factors = c("batch", "condition", "antigen", "hla", "serostatus", "cut_off_10", "infection")))

```



## Overview analysis of Cytof data

```{r}
n_cells(sce)
```

```{r fig.width= 10}
plotCounts(sce, color_by = "antigen")
```

## export n_cells to add to metadata for cut_off
```{r}
n_cells <- table(sample = sce$sample_id)
write.xlsx(n_cells, "n_cells.xlsx")
```

## iterate and add n_cells to table and metadata - clear environment and re-run
## FILTER sce to remove cut_off - cut_off = less than 0.01% (rounded-up) and/or less than 10 events

```{r}
sce_cutoff_10 <- filterSCE(sce, cut_off_10 == "yes")
```


```{r}
n_cells(sce_cutoff_10)
```


```{r fig.width= 10}
plotCounts(sce_cutoff_10, color_by = "hla")
```



```{r fig.width= 5}
plotNRS(sce_cutoff_10, features = type_markers(sce_cutoff_10), color_by = "condition")
```



## Clustering of fcs files with over 0.01% of CD8 (or more than 10 events) tetramer positive events  

Cluster FLOWSom analysis - max 10 clusters <br/>

For reproducibility set random seed
```{r}
set.seed(5678)
sce_cutoff_10 <- cluster(sce_cutoff_10, features = NULL, 
               xdim = 10, ydim = 10, maxK = 10, 
               verbose = FALSE, seed = 4321) 
```



### sce_all : Cluster of sce on type AND state markers; using features = NULL
Cluster FLOWSom analysis - max 10 clusters <br/> cluster on "type" markers for memory phenotyping




```{r fig.width=4}
pbMDS(sce_cutoff_10, shape_by = "antigen", size_by = TRUE, label_by = NULL)
```

## Dimension reduction - UMAP
```{r}
set.seed(4321)
sce_cutoff_10 <- runDR(sce_cutoff_10, dr = "UMAP", cells = 500, features = NULL)
```

```{r}
plotDR(sce_cutoff_10, "UMAP", color_by = "meta10")

```





## 28.11.2022 Group data in sce_all into three groups

## group_a - pre_omicron - only pre_omicron samples 
## group_b - spike only samples
## group_c - nonspike only 

```{r}
sce_cutoff_10_groupa <- filterSCE(sce_cutoff_10, infection == "pre_omicron")
```


```{r}
sce_cutoff_10_groupb <- filterSCE(sce_cutoff_10, antigen == "Spike")
```


```{r}
sce_cutoff_10_groupc <- filterSCE(sce_cutoff_10, antigen == "Non_Spike")
```

## Plots of UMAP based on FLowSOM clustering and DR on sce_all (all markers inc type and state)

```{r}
plotDR(sce_cutoff_10, "UMAP", color_by = "condition")

```


```{r}
plotDR(sce_cutoff_10_groupa, "UMAP", color_by = "condition")

```

```{r}
plotDR(sce_cutoff_10_groupb, "UMAP", color_by = "condition")

```

```{r}
plotDR(sce_cutoff_10_groupc, "UMAP", color_by = "condition")

```
## Heatmaps


```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_cutoff_10, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```


```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_cutoff_10_groupa, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_cutoff_10_groupb, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_cutoff_10_groupc, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

# Marker expression 

pre_omicron_spike_naive", "post_omicron_spike_naive", "pre_omicron_spike_conv", "post_omicron_spike_conv", "post_omicron_nonspike_naive", "pre_omicron_nonspike_conv", "post_omicron_nonspike_conv",  "pre_omicron_EBV", "post_omicron_EBV

```{r}
my_comparisons_groupa <- list(  c("pre_omicron_spike_naive", "pre_omicron_spike_previously_infected"), c("pre_omicron_spike_naive", "pre_omicron_nonspike_previously_infected"), c("pre_omicron_spike_naive", "pre_omicron_EBV"), c("pre_omicron_spike_previously_infected", "pre_omicron_nonspike_previously_infected"), c("pre_omicron_spike_previously_infected", "pre_omicron_EBV"), c("pre_omicron_nonspike_previousy_infected","pre_omicron_EBV" ))
```

```{r fig.height=8}
plotPbExprs(sce_cutoff_10_groupa, features=NULL, facet_by = "antigen", ncol = 3) + stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all." )
```

```{r fig.width=5}
plotPbExprs(sce_cutoff_10_groupa, features=NULL, facet_by = "antigen", ncol = 3) 
```

```{r fig.height= 8}
plotPbExprs(sce_cutoff_10_groupb, features=NULL, facet_by = "antigen", ncol = 3 ) + stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all." )
```


```{r fig.width=5}
plotPbExprs(sce_cutoff_10_groupb, features=NULL, facet_by = "antigen", ncol = 3) 
```

```{r fig.height=8}
plotPbExprs(sce_cutoff_10_groupc, features=NULL, facet_by = "antigen", ncol = 3) +  stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all." )
```



```{r fig.width=5}
plotPbExprs(sce_cutoff_10_groupc, features=NULL, facet_by = "antigen", ncol = 3) 
```

```{r}
sessionInfo()
```









